\LARGE \bf Planning with Attitude

### 2020-10-12 paper/main.tex
Addressed Zac's comments.

    Newton's method uses derivative information about a function to iteratively
    approximate its roots. For unconstrained systems, this method is highly effective,
    and can exhibit quadratic convergence. For constrained systems, the updated parameter
    can be projected back onto the feasible set at each iteration, but without the same
    convergence guarantees. For the constraints on $SO(3)$, Newton's method struggles to
    converge past a certain threshold due to this projection. By leveraging the
    quaternion calculus introduced, Newton's method can be modified to implicitly account
    for these constraints. To demonstrate this, we will examine Wahba's Problem. In 1965,
    Grace Wahba proposed the criterion for a least squares estimate of a spacecraft's
    attitude from vector measurements \cite{Wahba1965, markley2014fundamentals}. We will solve this
    problem using a standard nonlinear least squares method, as well as a method that
    exploits the true group structure of $SO(3)$ using the quaternion calculus presented
    here.

    \subsection{Methodology}
\todo{Let's keep our notation and language consistent for body vs. inertial frames. I propose ``body'' (B) and ``inertial'' or ``world'' (W), which is the standard in robotics.}
    Given known vectors in some inertial frame, ${}^N v_i$, and measurements of these
    vectors in some body fixed frame, ${}^B v_i$, our goal is to determine the relative
    rotation from the body to inertial frame ${}^N q^B$, expressed as a quaternion. We
    can define Wahba's loss function as the following: 
    \begin{align} \label{eq:wahba_loss}
        L &= \sum_i w_i \norm{{}^N v_i - {}^N A(q)^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
    \end{align} 
    \todo{Let's leave the $w_i$ off for clarity since you're not using it later.}
    where $r_i(q)$ is the residual vector.

    We can solve for ${}^N q^B$ using a nonlinear least squares method minimizing
    Wahba's loss function:
    \begin{mini*}
        {q}{ \norm{r(q)}_2^2 }{}{}
        \addConstraint { q\in SO(3).}
    \end{mini*}
    \todo{This should be $q \in \Q$ here, since unit quaternions are not isomorphic to $SO(3)$ (rotation matrices).}
    Following the typical approach for Newton's method, we minimize \eqref{eq:wahba_loss}
    by setting the gradient to zero:
    \begin{equation} \label{eq:newton_foc}
        \begin{aligned}
            \pdv{L}{q}^T &= \sum_i \pdv{r(q)}{q}^T r_i{q} := J^T r(q) = 0 \\
              &= \sum_i (-2 H^T R(q)^T R({}^B \hat{v}_i)^T r(q).
        \end{aligned}
    \end{equation}
    which can be obtained from the chain rule and \eqref{eq:quaternion_rotation}.

@@

    \begin{algorithm} 
    	\begin{algorithmic}[1]
    		\caption{Projected Gauss-Newton Method}\label{alg:pgn}
    		\State $k = 0$
    		\While{significant progress}
    		    \State $J = \pdv{r(q_k)}{ \partial q}$ 
    		    \State $ \delta q = -(J^TJ)^{-1}J^T r(q_k)$ 
    		  %  \State $q_{n+1} = \text{normalize}(q_n + q_{step})$ 
    		  \State $q_{k+1} = \Pi_{SO(3)}(q_k + \delta q)$ 
    		    \State $k = k + 1$
    		\EndWhile
    	\end{algorithmic}
    \end{algorithm}

    Alternatively, if we instead minimize with respect to a differential quaternion $\phi$,
    we adapt the algorithm by simply ``correcting'' our Jacobian $J$ using 
    \eqref{eq:quat_jacobian}:
    \begin{equation} 
        \bar{J} = \pdv{r(\mathbf{q} \otimes \varphi(\phi))}{\phi} = \pdv{r(q)}{q} G(q).
    \end{equation}
    We obtain our step---this time in the actual tangent space---as before:
    $\phi = (\bar{J}^T \bar{J})^{-1} \bar{J}^T)$. To obtain our next iterate, we ``add'' the
    step using the correct notion of composition for the group: 
    $\q_{k+1} = \q_k \otimes \varphi(\phi)$. This ``multiplicative'' Newton algorithm is 
    summarized in Algorithm \ref{alg:mgn}.

@@

    \subsection{Satellite Attitude Keep-Out}
    \todo{Let's add a little more detail here: How about saying the dynamics are \eqref{eq:rigid_body_dynamics} plus a few extra states to capture the bending modes. Also, let's just call the start tracker a camera or sensor.}
    A spacecraft with flexible appendages must perform a 150 degree slew while ensuring
    that a star tracker line-of-sight does not cross within 40 degrees of the sun. The
    attitudes that result in the star tracker pointing too close to the sun will be
    referred to as a keep-out zone. The flexible body spacecraft dynamics are based on
    the truncated hybrid coordinate model, with the three largest modes captured
    \cite{Tracy2020}. We can guarantee the star tracker body fixed vector ($^Br_{st}$)
    does not look within 40 degrees of the inertial sun vector ($^Nr_{sun}$) by
    formulating the following constraint:
    \todo{Let's use standard matrix notation here rather than the angle brackets, and let's also write the rotation correctly using $A(q)$ like in the other sections.}
    \begin{equation}
        \langle ^N q ^B {}^Br_{st} , ^Nr_{sun}\rangle \leq \cos(40 ^\circ).
    \end{equation}
    %Due to this constraint on the feasible attitudes, an eigen-axis slew trajectory planner may end up violating this keep-out zone \cite{markley2014fundamentals}. ALTRO can generate a feasible trajectory for this problem since it is able to reason about the nonlinear flexible spacecraft dynamics, actuator constraints, and the star tracker line of sight constraint.
        \begin{figure}
            \centering
            \includegraphics{figures/kevins_plots/attitude_slew_plot.tikz}
            \caption{Visualization of the flexible spacecraft slew with a keep-out zone. Attitude is parameterized with a Rodrigues parameter to visualize the trajectory in three dimensions. The constraint surface represents attitudes where the star tracker line-of-sight is within 40$^\circ$ of the sun. The unconstrained solution violates this constraint, while the constrained solution is able to avoid the keep out zone.}
        \end{figure}
\todo{Talk a little about the results in the plot here.}

>>>>

    Newton's method uses derivative information about a function to iteratively
    approximate its roots. For unconstrained systems, this method is highly effective,
    and can exhibit quadratic convergence. For constrained systems, the updated parameter
    can be projected back onto the feasible set at each iteration, but without the same
    convergence guarantees. For the constraints on $SO(3)$, Newton's method struggles to
    converge past a certain threshold due to this projection. By leveraging the
    quaternion calculus introduced, Newton's method can be modified to implicitly account
    for these constraints. To demonstrate this, we will examine Wahba's Problem. In 1965,
    Grace Wahba proposed the criterion for a least squares estimate of a spacecraft's
    attitude from vector measurements \cite{Wahba1965, markley2014fundamentals}. We will solve this
    problem using a modified version of Newton's method that
    exploits the true group structure of $SO(3)$ using the quaternion calculus presented
    here.

    \subsection{Methodology}
% \todo{Let's keep our notation and language consistent for body vs. inertial frames. I propose ``body'' (B) and ``inertial'' or ``world'' (W), which is the standard in robotics.}
    Given known vectors in some world frame, ${}^W v_i$, and measurements of these
    vectors in some body fixed frame, ${}^B v_i$, our goal is to determine the relative
    rotation from the body to world frame ${}^W q^B$, expressed as a quaternion. We
    can define Wahba's loss function as the following: 
    \begin{align} \label{eq:wahba_loss}
        L &= \sum_i \norm{{}^W v_i - {}^W A(q)^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
    \end{align} 
    % \todo{Let's leave the $w_i$ off for clarity since you're not using it later.}
    where $r_i(q)$ is the residual vector.

    We can solve for ${}^W q^B$ using a nonlinear least squares method minimizing
    Wahba's loss function:
    \begin{mini*}
        {q}{ \norm{r(q)}_2^2 }{}{}
        \addConstraint { q\in \Q.}
    \end{mini*}
    % \todo{This should be $q \in \Q$ here, since unit quaternions are not isomorphic to $SO(3)$ (rotation matrices).}
    Following the typical approach for Newton's method, we minimize \eqref{eq:wahba_loss}
    by setting the gradient to zero:
    \begin{equation} \label{eq:newton_foc}
        \begin{aligned}
            \pdv{L}{q}^T &= \sum_i \pdv{r(q)}{q}^T r_i{q} := J^T r(q) = 0 \\
              &= \sum_i (-2 H^T R(q)^T R({}^B \hat{v}_i)^T r(q).
        \end{aligned}
    \end{equation}
    which can be obtained from the chain rule and \eqref{eq:quaternion_rotation}.

@@

    % \begin{algorithm} 
    % 	\begin{algorithmic}[1]
    % 		\caption{Projected Gauss-Newton Method}\label{alg:pgn}
    % 		\State $k = 0$
    % 		\While{significant progress}
    % 		    \State $J = \pdv{r(q_k)}{ \partial q}$ 
    % 		    \State $ \delta q = -(J^TJ)^{-1}J^T r(q_k)$ 
    % 		  %  \State $q_{n+1} = \text{normalize}(q_n + q_{step})$ 
    % 		  \State $q_{k+1} = \Pi_{SO(3)}(q_k + \delta q)$ 
    % 		    \State $k = k + 1$
    % 		\EndWhile
    % 	\end{algorithmic}
    % \end{algorithm}

    Alternatively, if we instead minimize with respect to a differential quaternion $\phi$ \todo{Brian: is this supposed to be "differential rotation"?},
    we adapt the algorithm by simply ``correcting'' our Jacobian $J$ using 
    \eqref{eq:quat_jacobian}:
    \begin{equation} 
        \bar{J} = \pdv{r(\mathbf{q} \otimes \varphi(\phi))}{\phi} = \pdv{r(q)}{q} G(q).
    \end{equation}
    We obtain our step---this time in the actual tangent space---as before:
    $\phi = (\bar{J}^T \bar{J})^{-1} \bar{J}^T)$. To obtain our next iterate, we ``add'' the
    step using the correct notion of composition for the group: 
    $\q_{k+1} = \q_k \otimes \varphi(\phi)$. This ``multiplicative'' Newton algorithm is 
    summarized in Algorithm \ref{alg:mgn}.

@@

    \subsection{Satellite Attitude Keep-Out}
    \todo{Let's add a little more detail here: How about saying the dynamics are \eqref{eq:rigid_body_dynamics} plus a few extra states to capture the bending modes. Also, let's just call the start tracker a camera or sensor.}
    A spacecraft with flexible appendages must perform a 150 degree slew while ensuring
    that a star tracker line-of-sight does not cross within 40 degrees of the sun. The
    attitudes that result in the star tracker pointing too close to the sun will be
    referred to as a keep-out zone. The flexible body spacecraft dynamics are based on
    the truncated hybrid coordinate model, with the three largest modes captured
    \cite{Tracy2020}. We can guarantee the star tracker body fixed vector ($^Br_{st}$)
    does not look within 40 degrees of the inertial sun vector ($^Nr_{sun}$) by
    formulating the following constraint:
    % \todo{Let's use standard matrix notation here rather than the angle brackets, and let's also write the rotation correctly using $A(q)$ like in the other sections.}
    \begin{equation}
         [{}^W r_{sun}]^T[{}^W A(q) {}^B\, {}^Br_{st} ] \leq \cos(40 ^\circ).
    \end{equation}
    %Due to this constraint on the feasible attitudes, an eigen-axis slew trajectory planner may end up violating this keep-out zone \cite{markley2014fundamentals}. ALTRO can generate a feasible trajectory for this problem since it is able to reason about the nonlinear flexible spacecraft dynamics, actuator constraints, and the star tracker line of sight constraint.
        \begin{figure}
            \centering
            \includegraphics{figures/kevins_plots/attitude_slew_plot.tikz}
            \caption{Visualization of the flexible spacecraft slew with a keep-out zone. Attitude is parameterized with a Rodrigues parameter to visualize the trajectory in three dimensions. The constraint surface represents attitudes where the star tracker line-of-sight is within 40$^\circ$ of the sun. The unconstrained solution violates this constraint, while the constrained solution is able to avoid the keep out zone.}
            \label{fig:keepout}
        \end{figure}
% \todo{Talk a little about the results in the plot here.}
From Figure \ref{fig:keepout} we can see that without the pointing constraint present, the trajectory will pass through the keep-out zone, making the trajectory infeasible. With the keep-out zone constraint present, ALTRO is able to converge on a trajectory that .
